---
title: "R Notebook"
output: html_notebook
---

## Loading Libraries

```{r}
library(tidyr)
library(dplyr)
library(jsonlite)
```

## Load Data

```{r}
 MovieData <- read.csv('movies_metadata original.csv')
 # Some numeric data fields are stored as characters, so this must be changed
 str(MovieData)
```

## Data Analysis and Preprocesing

Convert Data columns to numeric

Modify Genres and production column from json string to single string specifying Genre and production house

```{r}
MovieData$popularity <- as.numeric(MovieData$popularity)
MovieData$vote_count <- as.numeric(MovieData$vote_count)
MovieData$runtime <- as.numeric(MovieData$runtime)
MovieData$budget <- as.numeric(MovieData$budget)
```

```{r}
MovieData <- unite(MovieData, title_tagline, original_title, tagline, sep = " - ")
MovieData <- select(MovieData, -"adult",-"id",-"imdb_id",-"original_language", -"overview", -"status",-"title", -"video", -"spoken_languages") %>%
   mutate(net_return = revenue - budget, ROI = net_return / budget * 100)
```

```{r}
MovieData$genres <- lapply(MovieData$genres, function(x) gsub(".*name.: '(.*)'.*","\\1",x))
MovieData$production_countries <- lapply(MovieData$production_countries, function(x) gsub(".*name.: '(.*)'.*","\\1",x))
MovieData$production_companies <- lapply(MovieData$production_companies, function(x) gsub(".*name.: '(.*)'.*","\\1",x))

```

```{r}
MovieData$release_date <- substr(MovieData$release_date, nchar(as.character(MovieData$release_date))-3, nchar(as.character(MovieData$release_date)))
MovieData$release_date <- as.numeric(MovieData$release_date)

```

```{r}
library(ggplot2)
MovieData %>% 
  unnest(genres) %>% 
  group_by(genres) %>% 
  summarise(revenue = sum(revenue)) %>% 
  ggplot(aes(x = genres, y = revenue)) +
  geom_bar(stat = "identity",fill = "blue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "Genres", y = "Revenue", title = "Revenue by Genres")

```

```{r}
MovieData %>% 
  unnest(release_date) %>% 
  group_by(release_date) %>% 
  summarise(revenue = sum(revenue)) %>% 
  ggplot(aes(x = release_date, y = revenue)) +
  geom_bar(stat = "identity",fill = "red") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "release_date", y = "Revenue", title = "Revenue by every year")
```

```{r}
library(tm)
dtm <- DocumentTermMatrix(MovieData$tagline_title)
dtm
```

```{r}
glimpse(MovieData)
```

Converting Production column to binary 1 for US Production house else foreign

```{r}
MovieData$production_countries <- as.character(MovieData$production_countries)
MovieData$production_countries <- ifelse(MovieData$production_countries == "United States of America", 1, 0)
```

selecting the columns that has only numerical values

```{r}
MovieData_n <- MovieData[,c(1,4,7,8,9,10,11)]
```

```{r}
MovieData_n <- MovieData_n[c("budget","popularity","release_date","runtime","vote_average","vote_count","revenue")]
```

```{r}
names(MovieData_n)

```

```{r}
# number of missing values in each column
colSums(is.na(MovieData_n))
# removing Missing values
MovieData_n <- na.omit(MovieData_n)
```

## Split data into Train and Test sets

```{r}
# split the data into training and testing data
set.seed(123)
train_sample <- sample(1:nrow(MovieData_n), 0.7*nrow(MovieData_n))
train_data <- MovieData_n[train_sample,]
test_data <- MovieData_n[-train_sample,]
```

## Linear Regression Model

```{r}
library(lattice)
library(ggplot2)
library(caret)
# Linear regression model
lm_model <- train(revenue ~ ., data = train_data, method = "lm", trControl = trainControl(method = "cv", number = 10))

# measure the model preformence
lm_pred <- predict(lm_model, test_data)
lm_mse <- sqrt(mean((test_data$revenue - lm_pred)^2))
lm_mse
```

```{r}
summary(lm_model)
```

## Logistic Regression

```{r}
log_model<- train(revenue ~ ., data = train_data, method = "glm", trControl = trainControl(method = "cv", number = 10))

# measure the model preformence
log_pred <- predict(log_model, test_data)
log_mse <- mean((test_data$revenue - log_pred)^2)
log_mse <- sqrt(log_mse)
log_mse
```

```{r}
summary(log_model)
```

## SVM - Linear

```{r}
library(e1071)
model_svm <- train(revenue ~ ., data = train_data, method = "svmLinear", trControl = trainControl(method = "cv", number = 10))
# measure the model preformence
pred_svm <- predict(model_svm, test_data)
mse_svm <- mean((test_data$revenue - pred_svm)^2)
mse_svm
```

```{r}
summary(model_svm)
```

## SVM - Radial

```{r}
model_svmr <- train(revenue ~ ., data = train_data, method = "svmRadial", trControl = trainControl(method = "cv", number = 10))
# measure the model preformence
pred_svmr <- predict(model_svmr, test_data)
mse_svmr <- mean((test_data$revenue - pred_svmr)^2)
mse_svmr
```

```{r}
summary(model_svmr)
```

## XGBoost

```{r}
model_xg <- train(revenue ~ ., data = train_data, method = "xgbTree", trControl = trainControl(method = "cv", number = 10))
# measure the model preformence
pred_xg <- predict(model_xg, test_data)
mse_xg <- mean((test_data$revenue - pred_xg)^2)
mse_xg
```

```{r}
model_xg
```

## Random Forest

```{r}
model_forest <-train(revenue ~ ., data = train_data, method = "rf", trControl = trainControl(method = "cv", number = 10))
# measure the model preformence
pred_forest <- predict(model_forest, test_data)
mse_forest <- mean((test_data$revenue - pred_forest)^2)
mse_forest

```

```{r}
summary(model_forest)
```

## Decision Tree

```{r}

#fit for decision tree model to predict revenue based on the budget, popularity, production_countries, runtime, vote_average, vote_count
model_tree <- train(revenue ~ ., data = train_data, method = "rpart", trControl = trainControl(method = "cv", number = 10))
# measure the model preformence
pred_tree <- predict(model_tree, test_data)
mse_tree <- sqrt(mean((test_data$revenue - pred_tree)^2))
mse_tree
```

```{r}
summary(model_tree)
```

### Split Training set into Train and Validation

```{r}
train_val_sample <- sample(1:nrow(train_data), 0.7*nrow(train_data))
train_data1 <- train_data[train_val_sample,]
val_data <- train_data[-train_val_sample,]
```

## Neural Networks

```{r}
library(tensorflow)
library(keras,warn.conflicts = FALSE)
```

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 10, activation = "relu", input_shape = 6) %>% 
  layer_dense(units = 15, activation = "relu")
  layer_dense(units = 1, activation = "sigmoid")
# compile the model with adam optimizer and mean square error loss function
model %>% compile(
  loss = "mean_squared_error",
  optimizer = "adam",
  metrics = "accuracy"
)
```

```{r}
history <- model %>% fit(
  as.matrix(train_data1[,1:6]), as.matrix(train_data1[,7]),
  epochs = 10,
  batch_size = 10,
  validation_data = list(as.matrix(val_data[,1:6]), as.matrix(val_data[,7])),
  verbose = 1
)   

model %>% evaluate( as.matrix(test_data[,1:6]),  as.matrix(test_data[,7]))

# measure the model preformence
pred <- predict(model, as.matrix(test_data[,1:6]))
mse_nn1 <- sqrt(mean((test_data$revenue - pred)^2))
mse_nn1
```

```{r}
plot(history)
```

```{r}
library(keras)
library(tfruns)
library(reticulate)
flags <- list(
  learning_rate=c(0.1,0.5,0.01),
                 units1=c(8, 16, 32),
                 units2=c(8, 16,32),
                 dropout=c(0.1, 0.3, 0.5),
                 batch_size=c(8, 16, 32) 
)
runs <- tuning_run('MovieData_Rev.R', sample = 0.01, flags = flags)
```

```{r}
training_run("MovieData_Rev.R")
```

### Neural Networks II

```{r}
model2 <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = 6) %>% 
  layer_dense(units = 32, activation = "relu")
  layer_dense(units = 1, activation = "sigmoid")
# compile the model with adam optimizer and mean square error loss function
model2 %>% compile(
  loss = "mean_squared_error",
  optimizer = "adam",
  metrics = "accuracy"
)
```

```{r}
history2 <- model2 %>% fit(
  as.matrix(train_data[,1:6]), as.matrix(train_data[,7]),
  epochs = 500,
  batch_size = 16,
  validation_split = 0.2,
  verbose = 1
)   

model2 %>% evaluate( as.matrix(test_data[,1:6]),  as.matrix(test_data[,7]))

# measure the model preformence
pred2 <- predict(model2, as.matrix(test_data[,1:6]))
mse_nn2 <- sqrt(mean((test_data$revenue - pred2)^2))
mse_nn2
```

```{r}
plot(history2)
```

## Summary of all the Models

```{r}
Movie_rev <- data.frame("Model" = c("Linear_reg","Log_reg","Decision Tree","SVM-Linear","SVM-Radial","XGBoost", "NN1","NN2"), "RMSE" = c( lm_mse ,log_mse, mse_svm,mse_svmr, mse_xg,mse_forest,mse_nn1,mse_nn2))

Movie_rev
```
